Modelling the gas kinematics of an atypical Lyman alpha emitting compact dwarf galaxy

### 2017-07-04 paper/paper.tex
paper.tex

We do not claim that our findings discard the existence of different
undiscovered models that could reproduce the observed features in \tol.
More involved models with multiple kinematic components can
in principle also be sucessful.
We focus on presenting two simple and succesful models that
manage, for the first time, to account for \tol's \lya line main
features.

@@

In order to map out this large parameter space, we randomly drew
$2500$ sets of parameters within a observationally realistic range,
based on the considerations of \citet{Laursen2013ApJ...766..124L}, 
yielding a large variety of single-, double- and triple-peaked
spectra. 
The full analysis of the the spectral features as well as
more details on the radiative transfer are presented by
\citet{Gronke2016}.

@@

\begin{equation}
\chi^2 = \sum_{i} \frac{({f}_i - \hat{f}_i)^2}{\sigma_i^2}, 
\label{eq:chi2}
\end{equation}
%
where $i$ iterates over velocity bins, $f_{i}$ is the observed flux,
$\sigma_i$ is the observed flux uncertainty and $\hat{f}_i$ is the
model estimated flux.
As we do not have an analytic expression for $\hat{f}$; we use the
results of the MonteCarlo to build histograms that play the role of
$\hat{f}$.

@@

\begin{figure*}
\begin{center}
\includegraphics[width=0.95\textwidth]{tree_result.pdf}
\caption{Classification tree example. 
The aim is to find the parameters that separate two
classes: low $\chi^2/d.o.f. < 35$ and high $\chi^2/d.o.f>35$. 
Each box describes the condition over the parameter of interest, the
level of sample impurity, the total number of samples and the value of the
number of samples in each class.  
In this example we randomly sample $80\%$ of the full data set of models to
start with $28$ models in the first class and
$1962$ models in the second class; the best way to increase the
probability to find a result with low $\chi^2/d.o.f.$ ($13$ models,
fourth bottom box, from left to the right)
is having the clump radial velocity, $v_{\infty,{\rm cl}}<157.0$ \kms;
the clump velocity dipersion, $\sigma_{{\rm cl}}>55.6$ \kms, and the
probability to have a \lya photon emitted in a clump, $P_{\rm cl}>0.683$.
The final classification is not perfect, showing that all other
parameters are also important but these three are the most relevant.  
\label{fig:tree}}
\end{center}
\end{figure*}

@@

Figure 1. summarizes our main finding.
Dots represent the observational data for \tol\ with the
overplot from our best fits from the analytical solution for the
multiphase model (thick line)  and the rotating homogeneous gas sphere
(thin line).

@@

In what follows we summarize the values of the kinematic parameters
that managed to explain \tol\'s \lya profile.

@@

We cannot discard that another region of parameter space
could also work.
It might be interesting to explore in the
future with a higher sampling of the parameter space once new 
observations  clarify the kinematic nature of \tol \footnote{In
  Section \S \ref{sec:conclusions} we describe  in more detail what
  kind of observations could be performed in the near future to
  improve our understanding of \tol's kinematics.} . 
Furthermore, considering the sparse sampling of parameter space and the results
from the random forest classifier, it is safer to consider that
the best values for $\sigma_{\rm cl}$, $v_{\infty {\rm cl}}$, $P_{\rm
  cl}$ are actually lower/upper/lower bounds, respectively.

@@

The prefered value for the rotational velocity can be explained as
follows. 
Lower rotational velocities than the favored value would produce a
double peaked line as the different doppler shifts produced on
different regions of the rotating sphere would not be large enough to
smear the double peaks into a single peak
\citep{GaravitoCamargo2014}. 
For the same reason, higher rotation velocities could produce a single
peak but the line would be broader than it is observed.
The fact that the velocity and optical depth priors were wide enough,
allows us to suggest that the current values for the spherical model
found by the MCMC are robust given the observational constraints.

@@

Which model has more
chances to be correct?  
We discard the rotation model by considering that the rotational
velocity of $\sim350$ \kms\ is too high for such a low
mass system. 
\tol's infered rotational velocity is larger than the
Milky Way's ($\sim200$\kms) while its baryonic mass is one order
of magnitude smaller. 
Furthermore, \tol's structural properties seem to be similar to known dwarf
elliptical galaxies. 
The sample of 7 dwarf ellipticals (dE) observed by
\cite{2003AJ....126.1794G} show similar properties as \tol. 
For instance, the dE galaxies IC 3344 and IC 794 have similar absolute $M_V$
magnitudes (-17.10 and -17.19) and sizes (0.90 kpc and 0.82 kpc) as
\tol. 
These dE galaxies have velocity dispersions of $31.1$ \kms and
$44.6$ \kms, respectively,  and their rotational velocities are lower
than $1$ \kms.  
This leads us to prefer the multiphase model as it provides 
a broad consistency with kinematic properties of dE galaxies.

Following the results of \cite{2011ApJ...726..108T}, who made a
detailed analysis for all kind of elliptical galaxies including dE galaxies, we
estimate a value for the dynamical mass using the constraints for
thevelocity dispersion, $\sigma$,  in a spherical system localized in a
region of size $r$

@@

These results would imply that \tol\ is part of the dE
class. 
However, this hypothesis opens up two puzzles.
First, dE galaxies are found either in clusters or groups of
galaxies; while \tol\ does not belong either to a group or a cluster
\citep{2001AJ....121..169F}. 
The second issue is that, unlike a dE, \tol\ hosts a bright and compact
startburst; imaging and spectroscopy studies suggest that the
startburst started less than 4 Myr ago, while the low surface
brightness regins are consistent with an age below one Gyr
\citep{2001AJ....121..169F}.

We suggest that new observational tests are needed to clarify the
kinematic nature of \tol\.
Integral field unit measurements resolving its spatial extent are up
to the task.  
\tol\ spans a region of $4$ arcseconds,
an instrument such as the Multi Unit Spectroscopic Explorer (MUSE)
\citep{2014Msngr.157...13B} with its nominal $0.2$ arcseconds spatial
sampling over a $1.0$ arcminute field in wide-field mode could provide
a coarse mapping of different ionization lines to infer a kinematic
map in the H$\alpha$ line, following the kind of work presented by
\citet{Herenz16} done with the Potsdam Multi-Aperture Spectrometer
(PMAS) \citep{PMAS} on nearby \lya emitting galaxies.

@@

In this particular case, we prefer the multiphase model becasue it has kinematic
conditions similar to those in dwarf ellipticals, while the
rotational model produces rotational velocities too high for a dwarf
galaxy.

@@

\section*{Acknowledgments}
We thank the referee for detailed comments and suggestions that
improved this paper. 
We thank J. M. Mas-Hesse for providing us with \tol's observational
\lya data \citep{mashesse03} in electronic format. This data was
used to prepare Figure 1.
J. E. F-R. ackwnoledges the support of J. L. Guerra in the last
writing stages of this paper.

>>>>

We do not claim that our findings discard the existence of different
undiscovered models that could reproduce the observed features in \tol.
More involved models with multiple kinematic components can
in principle also be successful.
We focus on presenting two simple and successful models that
manage, for the first time, to account for \tol's \lya line main
features.

@@

In order to map out this large parameter space, we randomly drew
$2500$ sets of parameters within a observationally realistic range,
based on the considerations of \citet{Laursen2013ApJ...766..124L}, 
yielding a large variety of single-, double- and triple-peaked
spectra. 
The full analysis of the spectral features as well as
more details on the radiative transfer are presented by
\citet{Gronke2016}.

@@

\begin{equation}
\chi^2 = \sum_{i} \frac{({f}_i - \hat{f}_i)^2}{\sigma_i^2}, 
\label{eq:chi2}
\end{equation}
%
where $i$ iterates over velocity bins, $f_{i}$ is the observed flux,
$\sigma_i$ is the observed flux uncertainty and $\hat{f}_i$ is the
model estimated flux.
As we do not have an analytic expression for $\hat{f}$; we use the
results of the Monte Carlo to build histograms that play the role of
$\hat{f}$.

@@

\begin{figure*}
\begin{center}
\includegraphics[width=0.95\textwidth]{tree_result.pdf}
\caption{Classification tree example. 
The aim is to find the parameters that separate two
classes: low $\chi^2/d.o.f. < 35$ and high $\chi^2/d.o.f>35$. 
Each box describes the condition over the parameter of interest, the
level of sample impurity, the total number of samples and the value of the
number of samples in each class.  
In this example we randomly sample $80\%$ of the full data set of models to
start with $28$ models in the first class and
$1962$ models in the second class; the best way to increase the
probability to find a result with low $\chi^2/d.o.f.$ ($13$ models,
fourth bottom box, from left to the right)
is having the clump radial velocity, $v_{\infty,{\rm cl}}<157.0$ \kms;
the clump velocity dispersion, $\sigma_{{\rm cl}}>55.6$ \kms, and the
probability to have a \lya photon emitted in a clump, $P_{\rm cl}>0.683$.
The final classification is not perfect, showing that all other
parameters are also important but these three are the most relevant.  
\label{fig:tree}}
\end{center}
\end{figure*}

@@

Figure 1. Summarizes our main finding.
Dots represent the observational data for \tol\ with the
overplot from our best fits from the analytical solution for the
multiphase model (thick line)  and the rotating homogeneous gas sphere
(thin line).

@@

In what follows we summarize the values of the kinematic parameters
that managed to explain \tol's \lya profile.

@@

We cannot discard that another region of parameter space
could also work.
It might be interesting to explore in the
future with a higher sampling of the parameter space once new 
observations  clarify the kinematic nature of \tol \footnote{In
  Section \S \ref{sec:conclusions} we describe  in more detail what
  kind of observations could be performed in the near future to
  improve our understanding of \tol's kinematics.}.
Furthermore, considering the sparse sampling of parameter space and the results
from the random forest classifier, it is safer to consider that
the best values for $\sigma_{\rm cl}$, $v_{\infty {\rm cl}}$, $P_{\rm
  cl}$ are actually lower/upper/lower bounds, respectively.

@@

The preferred value for the rotational velocity can be explained as
follows. 
Lower rotational velocities than the favored value would produce a
double peaked line as the different doppler shifts produced on
different regions of the rotating sphere would not be large enough to
smear the double peaks into a single peak
\citep{GaravitoCamargo2014}. 
For the same reason, higher rotation velocities could produce a single
peak but the line would be broader than it is observed.
The fact that the velocity and optical depth priors were wide enough,
allows us to suggest that the current values for the spherical model
found by the MCMC are robust given the observational constraints.

@@

Which model has more
chances to be correct?  
We discard the rotation model by considering that the rotational
velocity of $\sim350$ \kms\ is too high for such a low
mass system. 
\tol's inferred rotational velocity is larger than the
Milky Way's ($\sim200$\kms) while its baryonic mass is one order
of magnitude smaller. 
Furthermore, \tol's structural properties seem to be similar to known dwarf
elliptical galaxies. 
The sample of 7 dwarf ellipticals (dE) observed by
\cite{2003AJ....126.1794G} show similar properties as \tol. 
For instance, the dE galaxies IC 3344 and IC 794 have similar absolute $M_V$
magnitudes (-17.10 and -17.19) and sizes (0.90 kpc and 0.82 kpc) as
\tol. 
These dE galaxies have velocity dispersions of $31.1$ \kms and
$44.6$ \kms, respectively,  and their rotational velocities are lower
than $1$ \kms.  
This leads us to prefer the multiphase model as it provides 
a broad consistency with kinematic properties of dE galaxies.

Following the results of \cite{2011ApJ...726..108T}, who made a
detailed analysis for all kind of elliptical galaxies including dE galaxies, we
estimate a value for the dynamical mass using the constraints for
the velocity dispersion, $\sigma$,  in a spherical system localized in a
region of size $r$

@@

These results would imply that \tol\ is part of the dE
class. 
However, this hypothesis opens up two puzzles.
First, dE galaxies are found either in clusters or groups of
galaxies; while \tol\ does not belong either to a group or a cluster
\citep{2001AJ....121..169F}. 
The second issue is that, unlike a dE, \tol\ hosts a bright and compact
starburst; imaging and spectroscopy studies suggest that the
starburst started less than 4 Myr ago, while the low surface
brightness regins are consistent with an age below one Gyr
\citep{2001AJ....121..169F}.

We suggest that new observational tests are needed to clarify the
kinematic nature of \tol.
Integral field unit measurements resolving its spatial extent are up
to the task.  
\tol\ spans a region of $4$ arcseconds,
an instrument such as the Multi Unit Spectroscopic Explorer (MUSE)
\citep{2014Msngr.157...13B} with its nominal $0.2$ arcseconds spatial
sampling over a $1.0$ arcminute field in wide-field mode could provide
a coarse mapping of different ionization lines to infer a kinematic
map in the H$\alpha$ line, following the kind of work presented by
\citet{Herenz16} done with the Potsdam Multi-Aperture Spectrometer
(PMAS) \citep{PMAS} on nearby \lya emitting galaxies.

@@

In this particular case, we prefer the multiphase model because it has kinematic
conditions similar to those in dwarf ellipticals, while the
rotational model produces rotational velocities too high for a dwarf
galaxy.

@@

\section*{Acknowledgments}
We thank the referee for detailed comments and suggestions that
improved this paper. 
We thank J. M. Mas-Hesse for providing us with \tol's observational
\lya data \citep{mashesse03} in electronic format. This data was
used to prepare Figure 1.
J. E. F-R. acknowledges the support of J. L. Guerra in the last
writing stages of this paper.
